---
title: "eq_create_label"
author: "AH Uyekita"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(kableExtra)
library(msdr)
```

## Table of Contents

- [1. Introduction](#intro)  
- [2. Example 1](#example_1)
- [3. Example 2](#example_2)


## 1. Introduction {#intro}

This function compounds a new information to be displayed in a popup of Leaflet package. For this reason, the dataset should have `LOCATION`, `EQ_PRIMARY` and `DATE`.


## Example 1 {#example_1}

The next example only show the content of each popup of Leaflet map.

```{r, message=FALSE,warning=FALSE}
# Path to the raw data.
raw_data_path <- system.file("extdata", "signif.txt", package = "msdr")

# Loading the dataset of Earthquake.
df <- readr::read_delim(file = raw_data_path,      
                        delim = '\t',              
                        col_names = TRUE,          
                        progress = FALSE,           
                        col_types = readr::cols())

# Display the informaiton of each popup.
df %>% eq_create_label() %>% head() %>% kable()
```

## Example 2 {#example_2}

This example shows the complete case of plotting a popup with Location, Magnitude and date.

```{r, message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Display the informaiton of each popup.
df %>% 
       # Cleaninig the data.
       eq_clean_data() %>%
              # Creating new content.
              dplyr::mutate(popup_text = eq_create_label(.)) %>%
                     # Plotting new content inside of popup.
                     eq_map(annot_col = "popup_text")
```
