---
title: "eq_clean_data"
author: "AH Uyekita"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{eq_clean_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(kableExtra)
library(msdr)
```

## Table of Contents

- [1. Introduction](#intro)  
- [2. Loading the Data](#load_data)
- [3. Creating New Features](#new_features)
- [4. Conversion Process](#conversion)
- [5. Cleaning Process](#cleaning)
- [6. Example 1](#example_1)
- [7. Example 2](#example_2)

## 1. Introduction {#intro}


This function has two behaviours:

1) When you assign a file to load, and;

```
# Loading the 'signif.txt' file.
eq_clean_data(file_name = system.file("extdata", "signif.txt", package = "msdr"))
```

2) When you pipe a dataset already loaded.

```
# Pipe.
readr::read_delim("signif.txt",
                  delim = "\t") %>% eq_clean_data()
```

## 2. Loading the data {#load_data}

This function also loads the Earthquake database from NOAA.

```{r, message=FALSE,warning=FALSE, echo=FALSE}
# Path to the raw data.
raw_data_path <- system.file("extdata", "signif.txt", package = "msdr")

# Loading the dataset of Earthquake.
df <- readr::read_delim(file = raw_data_path,      
                        delim = '\t',              
                        col_names = TRUE,          
                        progress = FALSE,           
                        col_types = readr::cols())

# Printing the first 5 rows.
head(df) %>%
       select(I_D, YEAR, LOCATION_NAME, EQ_PRIMARY, TOTAL_DEATHS) %>% 
              kable()
```

As you can see, there are several observations with NA values.


## 3. Creating new features {#new_features}

The `eq_clean_data` creates the DATE variable binding the columns YEAR, MONTH, and DAY. All this using the [Lubridate][lubridate_url] package.

[lubridate_url]: https://lubridate.tidyverse.org

```{r, message=FALSE,warning=FALSE}
# Creating a new feature.
df <- df %>%
       mutate(DATE = lubridate::ymd(paste(df$YEAR,      # YEAR column
                                          df$MONTH,     # MONTH column
                                          df$DAY,       # DAY column
                                          sep = "/")))  # YYYY/MM/DD
```

## 4. Conversion Process {#conversion}

I have converted  the class of some features:

* `TOTAL_DEATHS` to numeric;
* `EQ_PRIMARY` to numeric;
* All `NA`'s of `TOTAL_DEATHS` in zeros.

## 5. Cleaning Process {#cleaning}

I have removed:

* All observations flagged as Tsunami, and;
* All observations with no Date.

## 6. Example 1 {#example_1}

How to load a `txt` file.

```{r, message=FALSE, warning=FALSE}

# Load the package
library(msdr)

# Define as file_name the txt file.
df <- eq_clean_data(file_name = raw_data_path)

# Dimensions of the loaded dataframe.
dim(df)
```


## 7. Example 2 {#example_2}

Piping a dataset to the `eq_clean_data`.

```{r, message=FALSE, warning=FALSE}

# Load the package
library(msdr)

# Piping a read_delim with eq_clean_data.
readr::read_delim(raw_data_path,
                  delim = "\t") %>%
       
              eq_clean_data() -> df

# Dimensions of the loaded dataframe.
dim(df)
```
