---
title: "eq_location_data"
author: "AH Uyekita"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(kableExtra)
library(msdr)
```

## Table of Contents

- [1. Introduction](#intro)  
- [2. Example](#example)

## 1. Introduction {#intro}

This function creates a new column with the earthquake `LOCATION`. The function `eq_clean_data` uses it behind the scenes, so it is not necessary to call this function after call `eq_clean_data`.

## Example {#example}

Piping a raw data to creates a LOCATION column.

```{r, message=FALSE,warning=FALSE}
# Path to the raw data.
raw_data_path <- system.file("extdata", "signif.txt", package = "msdr")

# Loading the dataset of Earthquake.
df <- readr::read_delim(file = raw_data_path,      
                        delim = '\t',              
                        col_names = TRUE,          
                        progress = FALSE,           
                        col_types = readr::cols())

# Printing some columns.
df %>%
       eq_location_clean() %>%
              # Selecting some features.
              select(YEAR,
                     COUNTRY,
                     LOCATION,
                     EQ_PRIMARY,
                     TOTAL_DEATHS) %>% 
                     # Filtering.
                     filter(YEAR > 1990 &
                            YEAR < 2019) %>%
                            # Show the first 10 rows.
                            head(10) %>%
                                   # Enhance table visualization.
                                   kable()
```

As you can see, the `LOCATION` column has only cities in Title Case mode. 
