---
title: "eq_map"
author: "AH Uyekita"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(kableExtra)
library(msdr)
```

## Table of Contents

- [1. Introduction](#intro)  
- [2. Example 1](#example_1)
- [3. Example 2](#example_2)

## 1. Introduction {#intro}

This function plots circles on OpenStreetMaps, which represents earthquakes. When you click in it appaers a popup, which shows a single information (that you define as `annot_col`).

## Example  {#example_1}

Plotting circles and popup Date of earthquake.

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Path to the raw data.
raw_data_path <- system.file("extdata", "signif.txt", package = "msdr")

# Loading the dataset of Earthquake.
df <- readr::read_delim(file = raw_data_path,      
                        delim = '\t',              
                        col_names = TRUE,          
                        progress = FALSE,           
                        col_types = readr::cols())

# Ploting all earthquakes from 1500 to 1700.
df %>% 
       # Converting data, cleaning and creating.
       eq_clean_data() %>%
              #Filtering.
              filter(YEAR > 1500 &
                     YEAR < 1700) %>%
                            # Adding circles and popups.
                            eq_map(annot_col = 'DATE')
```

The popups shows the `DATE` when you click the circle.


## Example 2 {#example_2}

Popup showing the Location of earthquake.

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Ploting all earthquakes from 1500 to 1700.
df %>% 
       # Converting data, cleaning and creating.
       eq_clean_data() %>%
              #Filtering.
              filter(YEAR > 1000 &
                     YEAR < 2019,
                     COUNTRY %in% 'CHILE') %>%
                            # Adding circles and popups.
                            eq_map(annot_col = 'LOCATION')

```

In this example I have selected the earthquakes in Chile. The function has autozoom to fit the circles in the view.
