---
title: "geom_timeline_label"
author: "AH Uyekita"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  
  options(warn=-1)
)
```


```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(kableExtra)
library(msdr)
library(ggplot2)
```

## Table of Contents

- [1. Introduction](#intro)  
- [2. Example](#example_1)

## 1. Introduction {#intro}

The `geom_timeline_label` is a complementary `geom` to aggregate notations to the `geom_timeline`. Using this `geom` you can add labels to the `n_max` most deadly earthquake. The configuration is almost the same of geom_timeline.

* mag: The magnitude of the earthquake;
* label: The Location of the earthquake;
* n_max: Number of earthquake to be highlighted using the notation;
* x axis: This is the temporal axis.
* y axis: Each county has your own line, it is not possible to mix countries in a single y axis.

## Example 1 {#example_1}

Let's plot the earthquake from 1900 to 2000, which occured in `CHINA` and `TURKEY`.

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Path to the raw data.
raw_data_path <- system.file("extdata", "signif.txt", package = "msdr")

# Loading the dataset of Earthquake.
df <- readr::read_delim(file = raw_data_path,      
                        delim = '\t',              
                        col_names = TRUE,          
                        progress = FALSE,           
                        col_types = readr::cols())

# Cleaning the data and filtering.
df %>% 
       eq_clean_data() %>%
       
              filter(COUNTRY %in% c('CHINA','TURKEY'),
                     YEAR >= 1900 &
                     YEAR <= 2000) %>%
              # Creating a ggplot object
              ggplot() + 
                     # Using the new Geom
                     geom_timeline(aes(x     = DATE,
                                       y     = COUNTRY,
                                       size  = EQ_PRIMARY,
                                       color = TOTAL_DEATHS)) +
                            
                            # Adding notation.
                            geom_timeline_label(aes(x = DATE,
                                                    y = COUNTRY,
                                                    mag = EQ_PRIMARY,
                                                    label = LOCATION,
                                                    n_max = 10)) + 
       
                                   # Adding theme.
                                   theme_msdr() + 
                                          # Editing the legends' titles 
                                          labs(size = "Richter scale value",
                                               color = "# deaths")
```

