---
title: "geom_timeline"
author: "AH Uyekita"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  
  options(warn=-1)
)
```


```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(kableExtra)
library(msdr)
library(ggplot2)
```

## Table of Contents

- [1. Introduction](#intro)  
- [2. Example 1](#example_1)
- [3. Example 2](#example_2)

## 1. Introduction {#intro}

The `geom_timeline` is a new `geom_*` of `ggplot2` package that aims to enhance the visualization of earthquake. This Geom has some configuration:

* size: The earthquakes as displayed as circles with different radius (according to the `EQ_PRIMARY`);
* color: This is based on the `TOTAL_DEATHS`;
* x axis: This is the temporal axis.
* y axis: Each county has your own line, it is not possible to mix countries in a single y axis.

## Example 1 {#example_1}

Let's plot the earthquake from 1000 to 2000, which occured in `JAPAN`.

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Path to the raw data.
raw_data_path <- system.file("extdata", "signif.txt", package = "msdr")

# Loading the dataset of Earthquake.
df <- readr::read_delim(file = raw_data_path,      
                        delim = '\t',              
                        col_names = TRUE,          
                        progress = FALSE,           
                        col_types = readr::cols())

# Cleaning the data and filtering.
df %>% 
       eq_clean_data() %>%
       
              filter(COUNTRY %in% 'JAPAN',
                     YEAR >= 1000 &
                     YEAR <= 2000) %>%
              # Creating a ggplot object
              ggplot() + 
                     # Using the new Geom
                     geom_timeline(aes(x     = DATE,
                                       y     = COUNTRY,
                                       size  = EQ_PRIMARY,
                                       color = TOTAL_DEATHS)) +
                            # Adding theme.
                            theme_msdr() + 
                                   # Editing the legends' titles 
                                   labs(size = "Richter scale value",
                                        color = "# deaths")
```

Most of earthquake records in Japan are concentrated between 1900 and 2000.

## Example 2 {#example_2}

The earthquake record of 2018. Simple comparison.

```{r}
# List of countries in Europe and "West Asia". This is not an exhaustive list.
eurasia <- c('SPAIN','GREECE','TURKEY','PORTUGAL','RUSSIA','FRANCE','MACEDONIA','BULGARIA',
             'ALBANIA','GEORGIA','ITALY','SLOVENIA','UK','CYPRUS','UKRAINE','CROATIA','AUSTRIA')

# Cleaninig data and filtering.
df %>% 
       eq_clean_data() %>%
              # Creating a new feature.
              mutate(CONTINENT = case_when(COUNTRY %in% eurasia ~ "EURASIA",
                                          !(COUNTRY %in% eurasia) ~ "WORLD")) %>%
                     # Filtering.
                     filter(YEAR >= 2018 &
                            YEAR <= 2019) %>%
              
                            # Creating a ggplot object
                            ggplot() + 
                                   # Using the new Geom
                                   geom_timeline(aes(x     = DATE,
                                                     y     = CONTINENT,
                                                     size  = EQ_PRIMARY,
                                                     color = TOTAL_DEATHS)) +
                                          # Adding theme.
                                          theme_msdr() + 
                                                 # Editing the legends' titles 
                                                 labs(size = "Richter scale value",
                                                      color = "# deaths")
```
