---
title: "theme_msdr"
author: "AH Uyekita"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  
  options(warn=-1)
)
```


```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(kableExtra)
library(msdr)
library(ggplot2)
```

## Table of Contents

- [1. Introduction](#intro)  
- [2. Example](#example)

## 1. Introduction {#intro}

I have decided to create a new theme because it is a good opportunity to doing it so I have used the default theme and I have changed some visuals. As you can see, I have made a great effort to make it be similar to the example posted in class notebooks.

## 2. Example {#example}

Adding theme to a regular plotting of earthquake.

*Without theme:*

```{r,message=FALSE,warning=FALSE}
# Path to the raw data.
raw_data_path <- system.file("extdata", "signif.txt", package = "msdr")

# Loading the dataset of Earthquake.
df <- readr::read_delim(file = raw_data_path,      
                        delim = '\t',              
                        col_names = TRUE,          
                        progress = FALSE,           
                        col_types = readr::cols())

# Cleaning the data and filtering.
df %>% 
       eq_clean_data() %>%
       
              filter(COUNTRY %in% 'USA',
                     YEAR >= 1900 &
                     YEAR <= 2000) %>%
              # Creating a ggplot object
              ggplot() + 
                     # Using the new Geom
                     geom_timeline(aes(x     = DATE,
                                       y     = COUNTRY,
                                       size  = EQ_PRIMARY,
                                       color = TOTAL_DEATHS)) -> earthquake_usa_1900_2000

# Plotting
earthquake_usa_1900_2000
```

*Adding theme:*

```{r,message=FALSE,warning=FALSE}
# Adding theme.
earthquake_usa_1900_2000 + theme_msdr()
```
