---
title: "Mastering Software Development in R"
author: "AH Uyekita"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  
  options(warn=-1)
)
```


```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(magrittr)
library(dplyr)
library(readr)
library(kableExtra)
library(msdr)
library(ggplot2)
```

## Table of Contents

- [1. Introduction](#intro)  
    - [1.1. Timeline Plots](#timeline)
    - [1.2. OpenStreet Maps](#openstreet)
- [2. MSDR Package](#msdr_package)
- [3. Usage](#usage)
    - [3.1. Loading Data](#loading_data)
    - [3.2. My first geom_timeline plot](#first_geom_timeline)
    - [3.3. Adding Annotations](#annotations)
    - [3.4. Earthquakes Info in OpenStreet maps](#openstreet_1)
    - [3.5. Earthquakes Popup in OpenStreet maps](#openstreet_2)

# 1. Introduction {#intro}

This collection of functions aims to enhance the visual experience of earthquakes. There are two kinds of functions:

1) Timeline based, and;
2) [OpenStreet][url_openstreet_site] Maps based.

[url_openstreet_site]: https://www.openstreetmap.org/

### Timeline Graphics {#timeline}

Draw a timeline point out each earthquake as a circle, the radius varies according to the magnitude (on Richter  scale). For each country (or category) there is a dedicated timeline. The user could use the function of annotation to highlight the strongest earthquakes from that database.

### OpenStreet Maps Plots {#openstreet}

This is an OpenStreet Map based plot, I have used the `LONGITUDE` and `LATITUDE` to plot circle as an earthquake, which the radius is according to the magnitude (in [Richter  scale][richter_scale]). By the using of the `eq_create_label`, it is possible to assign for each circle a popup to show information.

[richter_scale]: https://simple.wikipedia.org/wiki/Richter_scale

# 2. MSDR Package {#msdr_package}

The functions avaliable in this package:

* eq_clean_data
* eq_create_label
* eq_location_clean
* eq_map
* geom_timeline
* geom_timeline_label
* theme_msdr

A brief description of each of these functions.

### `eq_clean_data`

This function loads a given file_name and then performs the data cleaning. Undercover of this process these functions call the `eq_location_clean` to creates a new column called `LOCATION`.

### `eq_create_label`

Combines three columns to creates a new one with `HTML` structure, this is necessary because the Leaflet package requires the data to be displayed inside of the popup as HTML format.

### `eq_location_clean`

Adds the `LOCATION` column. The dataset must have the `LOCATION_NAME`. If not the function will not work properly.

### `eq_map`

Draw an OpenStreet Map and circles representing the earthquake's location. The popups show the date of the event. All this feature built over the [Leaflet][url_leaflet] package.

[url_leaflet]: https://rstudio.github.io/leaflet/

### `geom_timeline`

Plot a timeline based on magnitude (`EQ_PRIMARY`) and total deaths (`TOTAL_DEATHS`).

### `geom_timeline_label`

Given a plot of `geom_timeline`, this function annotates labels to the `n_max` earthquakes with the highest magnitude (`EQ_PRIMARY`).

### `theme_msdr`

A helper function to remove the background, grid, axis ticks, etc. Aims to increase the ink ratio of the plot.


There are two more functions, but these two has its works hidden.

### `GeomTimeline`

Creates all visuals to be plotted by the `geom_timeline`.

### `GeomTimelineLabel`

Creates all visuals to be plotted by the `geom_timeline_label`.

For further understanding, I kindly as you to visit the dedicated vignette of each function.

# 3. Usage {#usage}

This package is tailored to work with the [NOAA][noaa_url] (National Oceanic and Atmospheric Administration) [database of the earthquake][noaa_earthquake]. I have inserted in this package the entire database (downloaded in february/2019), you can access it creating the path and later loading it.

[noaa_url]: https://www.ngdc.noaa.gov
[noaa_earthquake]: https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1


```{r, message=FALSE,warning=FALSE}
# Path to the raw data.
raw_data_path <- system.file("extdata", "signif.txt", package = "msdr")
```

For the sake of this vignette, I have used this dataset to perform the desmonstrations.

## 3.1. Loading the NOAA Database {#loading_data}

The loading process could be done using the `eq_clean_data`.

```{r, message=FALSE,warning=FALSE}
# Loading the data.
df_clean <- eq_clean_data(file_name = raw_data_path)
```

After that, the `df_clean` has LOCATION column, LONGITUDE, LATITUDE, etc. converted to the proper class type.


```{r,echo=FALSE}
# Subsetting the df_clean. Showing only data of interest.
df_clean %>% dplyr::select(DATE,
                           COUNTRY, 
                           LOCATION,
                           EQ_PRIMARY,
                           TOTAL_DEATHS) %>% head(10) %>% kableExtra::kable()
```

The table above is an example of the output of the `eq_clean_data`. Bear in mind, I have only converted and cleaned the columns of interest (`LATITUDE`, `LONGITUDE`, `DATE`, `COUNTRY`, `LOCATION`, `EQ_PRIMARY`, and `TOTAL_DEATHS`).

_ATTENTION:_ The `eq_location_clean` works internally of the `eq_clean_data`, as a step of data manipulation.

## 3.2. Creating my first geom_timeline {#first_geom_timeline}

This is a straightforward application of the `geom_timeline` function. Before any plot, I have chosen Asia as my interest in analysis.

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Subsetting the df_clean to select some countries in Asia.
df_asia <- df_clean %>% 
       dplyr::filter(COUNTRY %in% c("INDONESIA","THAILAND", "MYANMAR (BURMA)", "JAPAN"),
                     YEAR > 2000 & YEAR <= 2019)

# Creaing a ggplot object using the geom_timeline.
df_asia %>%
       ggplot2::ggplot() +
              # Defining the aes.
              geom_timeline(aes(x = DATE,
                                y = COUNTRY,
                                size = EQ_PRIMARY,
                                color = TOTAL_DEATHS)) -> tl_asia

# Plotting
tl_asia
```

To enhance the visuals, I have created the `theme_msdr`, which removes the background, grids, etc..

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Plotting with tailored theme.
tl_asia + msdr::theme_msdr() -> tl_asia_w_theme

tl_asia_w_theme
```

You can go also further editing the general aspects of the plot using the regular functions of `ggplot2` package. In the example below I have edited the legends' titles.

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Plotting with tailored theme.
tl_asia_w_theme + 
       msdr::theme_msdr() + 
              # Editing the legends' titles 
              labs(color = "# deaths",
                   size = "Richter scale value") -> tl_asia_w_theme_labs

tl_asia_w_theme_labs
```             

## 3.3. Adding annotations {#annotations}

In a way to highlight the strongest events in the given dataframe of analysis. The `geom_timeline_label` is a complementary function to the `geom_timeline`. It will annotate which earthquake from the plot was in the top five (default but you can edit it defining a value to `n_max`).

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Plotting with tailored theme.
tl_asia_w_theme_labs +
       # Adding annotations
       geom_timeline_label(aes(x = DATE,
                               label = LOCATION,
                               y = COUNTRY,
                               mag = EQ_PRIMARY,
                               n_max = 10)) -> tl_asia_w_theme_labs_annot

tl_asia_w_theme_labs_annot
``` 

## 3.4. Earthquakes Info in OpenStreet maps {#openstreet_1}

The `msdr` package has two features to be used in OpenStreet maps, the first one is a straightforward way to plot the earthquakes, and a single content in a popup.

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Using the cleaned dataset of asia. Plotting DATE inside of the popups.
df_asia %>% 
       eq_map(annot_col = 'DATE')
```

In the example above, I have plotted the `DATE` of each earthquake. If you are going to use numeric variables, I kindly ask you to use `as.character` to convert the information in character.

Example using a native non-character feature.

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Using the cleaned dataset of asia. Plotting the TOTAL_DEATHS inside of the popups.
df_asia %>% 
       eq_map(annot_col = 'as.character(TOTAL_DEATHS)')
```

## 3.5. Earthquakes Popup in OpenStreet maps {#openstreet_2}

This is a step further of the simple `eq_map` function. The `eq_create_label` combine many columns into one column to plot it inside of a popup of OpenStreet maps.

```{r,message=FALSE,warning=FALSE,out.width="100%",fig.height=3.5}
# Creating a new data.
df_america <- df_clean %>% dplyr::filter(COUNTRY %in% c('USA','MEXICO','CANADA'),
                                         YEAR > 1990 & YEAR < 2019)

# Creating a complex texts using the eq_create_label.
df_america %>% 
       dplyr::mutate(popup_text = eq_create_label(.)) %>% 
              eq_map(annot_col = 'popup_text')
```
